---
title: "Essentials of Data Science bootcamp assignment 1"
author: "Kay Hiranaka"
date: "February 6, 2016"
output: html_document
---

<!-- ignore comment -->
## Assignment 1 : Grade Calculator##

For the first assignment, I calculated grades for a physics course. This course has two parts: lecture and lab. The lab part is divided into two sections (groups). Separate data tables were supplied for each lab section and lecture. The lecture is worth 85% of the final grade, and the lab is worth 15%. I calculated grades for lab and lecture separately, then combined them by the student ID number. 

### Lecture Grade ###
The lecture data table contains grades for two midterm and a final exams. The final is worth 45% of the lecture grade, the lowest midterm is worth 20%, and the highest midterm is 35%. The code chunk below shows how I computed lecture grades.

First I read the lecture grade sheet.
```{r, message=F }
library(xlsx)
library(dplyr)

lec <- read.xlsx('project1_data/lecture.xlsx',1)

lecw <- data.frame(ID = 1:32, MT1 = lec$MT1, MT2 = lec$MT2, final = lec$Final) #working data frame
```

Then, I calculate midterm grades based on how they did on each midterm.
```{r}
mtfrac1 <- 0.2*lecw[lecw$MT1 <= lecw$MT2, 'MT1'] + 0.35*lecw[lecw$MT1 <= lecw$MT2, 'MT2'] #midterm grade if midterm1<=midterm2
mtfrac2 <- 0.2*lecw[lecw$MT2 < lecw$MT1, 'MT2'] + 0.35*lecw[lecw$MT2 < lecw$MT1, 'MT1'] # midterm grade if midterm1>midterm2

```

I sort ID, final score, and midterm scores by whether they did better on midterm 1 or midterm 2.
```{r}
id <- c(lecw[lecw$MT1 <= lecw$MT2,'ID'],lecw[lecw$MT1 > lecw$MT2,'ID'])
final <- c(lecw[lecw$MT1 <= lecw$MT2,'final'],lecw[lecw$MT1 > lecw$MT2,'final'])
mt1 <- c(MT1 = lecw[lecw$MT1 <= lecw$MT2,'MT1'],MT1 = lecw[lecw$MT1 > lecw$MT2,'MT1']) #lowest midterm
mt2 <- c(MT2 = lecw[lecw$MT1 <= lecw$MT2,'MT2'],MT2 = lecw[lecw$MT1 > lecw$MT2,'MT2']) #highest midterm

```

Make a new sorted data frame and add midterm grade.
```{r}
lecww <- data.frame(ID=id,MT1=mt1,MT2=mt2,final=final)
MTfrac = c(mtfrac1,mtfrac2)
lecww <- mutate(lecww, MTfrac = c(mtfrac1,mtfrac2))

```

Finally, calculate overall lecture grade and add it to the data frame.
```{r}
lecg <- MTfrac + final*0.45 #grade for lecture
lecww <- mutate(lecww, lecgrade = MTfrac + final*0.45)
lecww <- lecww[order(lecww$ID),] # sort by ID
```


### Lab Grade ###
The lab data tables contain grades for 12 labs and quizes 1 & 2. Each quiz is worth 10% of the lab grade, and 10 best labs are worth 80%. 2 lowest labs are dropped. Here is my code chunk for lab section 12. An excel table was supplied for this section.

#### Section 12 ####
For section 12, I read the xlsx file and assign 0 to all NULL elements.
```{r}
library(stringr)
#calculate lab12 grade
lab12 <- read.xlsx('project1_data/Fall2015Lab120_Sec 1L12.xlsx',1,startRow = 8,endRow = 20)
lab12[is.na(lab12)] <- 0 # fill in NULL
lab12$X.FINAL.LAB.GRADE.AND.COMMENT <- NULL #drop column
```

Then calculate quiz grade based on quiz 1 and 2.
```{r}
quizfrac <- 0.1*5*(lab12$QUIZ.2+lab12$QUIZ.1) #calculate quiz grade
lab12 <- mutate(lab12, quizfrac=quizfrac)
```


```{r}

lab12w <- lab12 %>%
  mutate(Student.s.Name = str_replace(Student.s.Name, 's','') )
lab12w$QUIZ.1 <- NULL #drop column
lab12w$QUIZ.2 <- NULL #drop column
lab12w$Student.s.Name <- NULL #drop more column

lab12w <- t(apply(lab12w,1,sort)) #sort remaining rows (only labs)
lab12w <- lab12w[,-(1:2)] #drop two lowest labs

labfrac <- rowSums(lab12w)/10*0.8 #calculate lab grade
```


Similarly, I calculated lab grades for section 9. Data was provided in an ODS table.

```{r}
library(readODS)
#calculate lab9 grade
lab09 <- read.ods('project1_data/Fall2015Lab120section 1L09.ods',1)
lab09 <- lab09[-c(1:7),] #refine table
lab09 <- lab09[-c(22:41),] #patience
lab09 <- lab09[,-16] #getting there
lab09[1,14] <- 'Quiz1'
lab09[1,15] <- 'Quiz2'
names(lab09) <- lab09[1,]
lab09 <- lab09[-1,]

lab09w <- lab09
quizfr <- 0.1*5*(as.numeric(lab09w[,14])+as.numeric(lab09w[,15])) #calculate quiz grade
lab9 <- data.frame(ID=1:20,quizfrac=quizfr)

lab09w <- lab09w[,-c(14:15)] #drop quiz columns
lab09w <- lab09w[,-1] #and ID

lab09ww <- lab09[,1:13] 
lab09ww <- lab09ww[,-1]
lab09ww[lab09ww==100] <- 999 #strings are not behaving (they think 100 is small). I'm giving '100's a special treatment.
lab09ww <- t(apply(lab09ww,1,sort)) #now sort rows
lab09ww[lab09ww==999] <- 100 #put back 100
lab09ww <- lab09ww[,-(1:2)] #drop two lowest lab grades
lab09ww[lab09ww==''] <- 0 #replace blank elements with 0

lab09www <- as.data.frame(lab09ww, stringsAsFactors=FALSE) #here convert strings into numeric
for (column in 1:10) {
  lab09www[, column] <- as.numeric(lab09www[, column])
}
labfr <- rowSums(lab09www)/10*0.8 #calculate lab grades
lab09www <- mutate(lab09www,labfrac=labfr)

quizf <- c(quizfr,quizfrac) #combine quiz grades from two sections
labf <- c(labfr,labfrac) #combine lab grades
labg <- quizf+labf #add them up
grade <- lecg*0.85 + labg*0.15 #finally calculate final grades!
final_grade <- data.frame(ID = 1:32, lecgrade=lecg,quizfrace=quizf, 
                          labfrac=labf, labgrade=quizf+labf, finalgrade=grade)

```

Here is my final output. This table contains lab grade, lecture grade, and final grade for each student.

```{r}
read.csv('grade_table.csv')
```


